% ============================================================
% references.tex — Inline bibliography (no BibTeX needed)
% ============================================================

\begin{thebibliography}{35}

\bibitem{fjell2010structural}
A.~M. Fjell and K.~B. Walhovd,
``Structural brain changes in aging: courses, causes and cognitive consequences,''
\textit{Reviews in the Neurosciences}, vol.~21, no.~3, pp.~187--221, 2010.

\bibitem{salat2004thinning}
D.~H. Salat, R.~L. Buckner, A.~Z. Snyder, D.~N. Greve, R.~S.~R. Desikan, E.~Busa, J.~C. Morris, A.~M. Dale, and B.~Fischl,
``Thinning of the cerebral cortex in aging,''
\textit{Cerebral Cortex}, vol.~14, no.~7, pp.~721--730, 2004.

\bibitem{storsve2014differential}
A.~B. Storsve, A.~M. Fjell, C.~K. Tamnes, L.~T. Westlye, K.~Overbye, H.~W. Aasland, and K.~B. Walhovd,
``Differential longitudinal changes in cortical thickness, surface area and volume across the adult life span,''
\textit{Journal of Neuroscience}, vol.~34, no.~21, pp.~7488--7498, 2014.

\bibitem{raz2005regional}
N.~Raz, U.~Lindenberger, K.~M. Rodrigue, K.~M. Kennedy, D.~Head, A.~Williamson, C.~Dahle, D.~Gerstorf, and J.~D. Acker,
``Regional brain changes in aging healthy adults: general trends, individual differences and modifiers,''
\textit{Cerebral Cortex}, vol.~15, no.~11, pp.~1676--1689, 2005.

\bibitem{deary2010neuroscience}
I.~J. Deary, L.~Penke, and W.~Johnson,
``The neuroscience of human intelligence differences,''
\textit{Nature Reviews Neuroscience}, vol.~11, no.~3, pp.~201--211, 2010.

\bibitem{riccelli2017surface}
R.~Riccelli, N.~Toschi, S.~Nigro, A.~Terracciano, and L.~Passamonti,
``Surface-based morphometry reveals the neuroanatomical basis of the five-factor model of personality,''
\textit{Social Cognitive and Affective Neuroscience}, vol.~12, no.~4, pp.~671--684, 2017.

\bibitem{sexton2014poor}
C.~E. Sexton, A.~B. Storsve, K.~B. Walhovd, H.~Johansen-Berg, and A.~M. Fjell,
``Poor sleep quality is associated with increased cortical atrophy in community-dwelling adults,''
\textit{Neurology}, vol.~83, no.~11, pp.~967--973, 2014.

\bibitem{pietschnig2015meta}
J.~Pietschnig, L.~Penke, J.~M. Wicherts, M.~Zeiler, and M.~Voracek,
``Meta-analysis of associations between human brain volume and intelligence differences,''
\textit{Neuroscience \& Biobehavioral Reviews}, vol.~57, pp.~411--432, 2015.

\bibitem{colom2009human}
R.~Colom, R.~J. Haier, K.~Head, J.~{\'A}lvarez-Linera, M.~{\'A}. Quiroga, P.~C. Shih, and R.~E. Jung,
``Gray matter correlates of fluid, crystallized, and spatial intelligence: testing the P-FIT model,''
\textit{Intelligence}, vol.~37, no.~2, pp.~124--135, 2009.

\bibitem{jung2007pfit}
R.~E. Jung and R.~J. Haier,
``The Parieto-Frontal Integration Theory (P-FIT) of intelligence: converging neuroimaging evidence,''
\textit{Behavioral and Brain Sciences}, vol.~30, no.~2, pp.~135--154, 2007.

\bibitem{bjornebekk2013neuronal}
A.~Bj{\o}rnebekk, A.~M. Fjell, K.~B. Walhovd, H.~Grydeland, S.~Torgersen, and L.~T. Westlye,
``Neuronal correlates of the five factor model (FFM) of human personality: multimodal imaging in a large healthy sample,''
\textit{NeuroImage}, vol.~65, pp.~194--208, 2013.

\bibitem{jackson2011cortical}
J.~Jackson, D.~A. Balota, and D.~Head,
``Exploring the relationship between personality and regional brain structure in healthy aging,''
\textit{Neurobiology of Aging}, vol.~32, no.~12, pp.~2162--2171, 2011.

\bibitem{lupien2009effects}
S.~J. Lupien, B.~S. McEwen, M.~R. Gunnar, and C.~Heim,
``Effects of stress throughout the lifespan on the brain, behaviour and cognition,''
\textit{Nature Reviews Neuroscience}, vol.~10, no.~6, pp.~434--445, 2009.

\bibitem{vanessen2013wuminn}
D.~C. Van~Essen, S.~M. Smith, D.~M. Barch, T.~E.~J. Behrens, E.~Yacoub, and K.~Ugurbil,
``The WU-Minn Human Connectome Project: an overview,''
\textit{NeuroImage}, vol.~80, pp.~62--79, 2013.

\bibitem{buysse1989pittsburgh}
D.~J. Buysse, C.~F. Reynolds, T.~H. Monk, S.~R. Berman, and D.~J. Kupfer,
``The Pittsburgh Sleep Quality Index: a new instrument for psychiatric practice and research,''
\textit{Psychiatry Research}, vol.~28, no.~2, pp.~193--213, 1989.

\bibitem{costa1992neo}
P.~T. Costa and R.~R. McCrae,
\textit{Revised NEO Personality Inventory (NEO-PI-R) and NEO Five-Factor Inventory (NEO-FFI) Professional Manual}.
Odessa, FL: Psychological Assessment Resources, 1992.

\bibitem{fischl2012freesurfer}
B.~Fischl,
``FreeSurfer,''
\textit{NeuroImage}, vol.~62, no.~2, pp.~774--781, 2012.

\bibitem{desikan2006automated}
R.~S. Desikan, F.~S{\'e}gonne, B.~Fischl, B.~T. Quinn, B.~C. Dickerson, D.~Blacker, R.~L. Buckner, A.~M. Dale, R.~P. Maguire, B.~T. Hyman, M.~S. Albert, and R.~J. Killiany,
``An automated labeling system for subdividing the human cerebral cortex on MRI scans into gyral based regions of interest,''
\textit{NeuroImage}, vol.~31, no.~3, pp.~968--980, 2006.

\bibitem{glasser2016multimodal}
M.~F. Glasser, T.~S. Coalson, E.~C. Robinson, C.~D. Hacker, J.~Harwell, E.~Yacoub, K.~Ugurbil, J.~Andersson, C.~F. Beckmann, M.~Jenkinson, S.~M. Smith, and D.~C. Van~Essen,
``A multi-modal parcellation of human cerebral cortex,''
\textit{Nature}, vol.~536, no.~7615, pp.~171--178, 2016.

\bibitem{bookheimer2019aging}
S.~Y. Bookheimer, D.~H. Salat, M.~Terpstra, B.~M. Ances, D.~M. Barch, R.~L. Buckner, G.~C. Burgess, S.~W. Curtiss, M.~Diaz-Santos, J.~S. Elam, and others,
``The Lifespan Human Connectome Project in Aging: an overview,''
\textit{NeuroImage}, vol.~185, pp.~335--348, 2019.

\bibitem{benjamini1995controlling}
Y.~Benjamini and Y.~Hochberg,
``Controlling the false discovery rate: a practical and powerful approach to multiple testing,''
\textit{Journal of the Royal Statistical Society: Series B}, vol.~57, no.~1, pp.~289--300, 1995.

\bibitem{cohen1988statistical}
J.~Cohen,
\textit{Statistical Power Analysis for the Behavioral Sciences}, 2nd~ed.
Hillsdale, NJ: Lawrence Erlbaum Associates, 1988.

\bibitem{krause2017sleep}
A.~J. Krause, E.~B. Simon, B.~A. Mander, S.~M. Greer, J.~M. Saletin, A.~N. Goldstein-Piekarski, and M.~P. Walker,
``The sleep-deprived human brain,''
\textit{Nature Reviews Neuroscience}, vol.~18, no.~7, pp.~404--418, 2017.

\bibitem{branger2016relationships}
P.~Branger, E.~M. Arenaza-Urquijo, C.~Tomadesso, F.~M{\'e}zenge, C.~Andr{\'e}, R.~de~Flores, J.~Mutlu, V.~de~La~Sayette, F.~Eustache, G.~Ch{\'e}telat, and G.~Rauchs,
``Relationships between sleep quality and brain volume, metabolism, and amyloid deposition in late adulthood,''
\textit{Neurobiology of Aging}, vol.~41, pp.~107--114, 2016.

\bibitem{spira2016self}
A.~P. Spira, C.~E. Gonzalez, V.~K. Venkatraman, M.~N. Wu, J.~Pacheco, E.~M. Simonsick, L.~Ferrucci, and S.~M. Resnick,
``Sleep duration and subsequent cortical thinning in cognitively normal older adults,''
\textit{Sleep}, vol.~39, no.~5, pp.~1121--1128, 2016.

\bibitem{fjell2023poor}
A.~M. Fjell, {\O}.~S{\o}rensen, I.~K. Amlien, D.~Bartr{\'e}s-Faz, D.~M. Bros, N.~Buchmann, and others,
``Self-reported sleep relates to hippocampal atrophy across the adult lifespan: results from the Lifebrain consortium,''
\textit{Sleep}, vol.~43, no.~5, zsz280, 2020.

\bibitem{grasby2020genetic}
K.~L. Grasby, N.~Jahanshad, J.~N. Painter, L.~Colodro-Conde, J.~Bralten, D.~P. Hibar, and others,
``The genetic architecture of the human cerebral cortex,''
\textit{Science}, vol.~367, no.~6484, eaay6690, 2020.

\bibitem{vuoksimaa2016cortical}
E.~Vuoksimaa, M.~S. Panizzon, C.~H. Chen, M.~Fiecas, L.~T. Eyler, C.~Fennema-Notestine, and others,
``Is bigger always better? The importance of cortical configuration with respect to cognitive ability,''
\textit{NeuroImage}, vol.~129, pp.~356--366, 2016.

\bibitem{servaas2015neuroticism}
M.~N. Servaas, J.~van~der~Velde, S.~G. Costafreda, P.~Horton, J.~Ormel, H.~Riese, and A.~Aleman,
``Neuroticism and the brain: a quantitative meta-analysis of neuroimaging studies investigating emotion processing,''
\textit{Neuroscience \& Biobehavioral Reviews}, vol.~37, no.~8, pp.~1518--1529, 2013.

\bibitem{hogstrom2013structure}
L.~J. Hogstrom, L.~T. Westlye, K.~B. Walhovd, and A.~M. Fjell,
``The structure of the cerebral cortex across adult life: age-related patterns of surface area, thickness, and gyrification,''
\textit{Cerebral Cortex}, vol.~23, no.~11, pp.~2521--2530, 2013.

\bibitem{lemaitre2012normal}
H.~Lemaitre, A.~L. Goldman, F.~Sambataro, B.~A. Verchinski, A.~Meyer-Lindenberg, D.~R. Weinberger, and V.~S. Mattay,
``Normal age-related brain morphometric changes: nonuniformity across cortical thickness, surface area and gray matter volume?''
\textit{Neurobiology of Aging}, vol.~33, no.~3, pp.~617.e1--617.e9, 2012.

\bibitem{mcewen2006protective}
B.~S. McEwen,
``Protective and damaging effects of stress mediators: central role of the brain,''
\textit{Dialogues in Clinical Neuroscience}, vol.~8, no.~4, pp.~367--381, 2006.

\bibitem{ottaviani2016physiological}
C.~Ottaviani, J.~F. Thayer, B.~Verkuil, A.~Lonigro, B.~Medea, A.~Couyoumdjian, and J.~F. Brosschot,
``Physiological concomitants of perseverative cognition: a systematic review and meta-analysis,''
\textit{Psychological Bulletin}, vol.~142, no.~3, pp.~231--259, 2016.

\bibitem{lim2013sleep}
A.~S.~P. Lim, M.~Kowgier, L.~Yu, A.~S. Buchman, and D.~A. Bennett,
``Sleep fragmentation and the risk of incident Alzheimer's disease and cognitive decline in older persons,''
\textit{Sleep}, vol.~36, no.~7, pp.~1027--1032, 2013.

\bibitem{sutin2011personality}
A.~R. Sutin, A.~Terracciano, M.~H. Kitner-Triolo, M.~Uda, D.~Schlessinger, and A.~B. Zonderman,
``Personality and metabolic syndrome,''
\textit{Age}, vol.~33, no.~4, pp.~573--582, 2011.

\end{thebibliography}
